---
title: "EXPLORE NEON Workshop"
output: md_document
date: 11/09/18
---

Task: Compare outputs of NEON canopy height model to actual data.

```{r setup, message = FALSE}

options(stringsAsFactors = F)

#load libraries
library(neonUtilities)
library(geoNEON)
library(raster)
library(ggplot2)
library(tidyverse) #this is important for data processing!

#make default plot white on white background
th <- theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
theme_set(th)


proj_dir <- "/Users/lilaleatherman/Documents/current projects/github/NEON_Explore/LeathermanLST_ExploreNEONData"

data_dir <- "/Users/lilaleatherman/Documents/current projects/github/neon_data_workshop/data"


```

```{r load veg structure data}

vegmap <- read.delim(paste0(data_dir, "/filesToStack10098/stackedFiles/vst_mappingandtagging.csv"), sep = ",")

vegind <- read.delim(paste0(data_dir, "/filesToStack10098/stackedFiles/vst_apparentindividual.csv"), sep = ",")

```

geoNEON provides functionality for spatial data around NEON data. 
For example, mapping and tagging data locations are relative to a named point, but don't have absolute location data themselves. geoNEON allows you to calculate the locations.

based on namedLocation, pointID, stemDistance, stemAzimuth - calculates the location by querying API

```{r look at stem diameter in plots}

#get lat / long from NEON
vegmap <- geoNEON::def.calc.geo.os(vegmap, "vst_mappingandtagging")

#join locations with stem diameter
veg <- merge(vegmap, vegind, by = c("individualID", "namedLocation", "domainID", "siteID", "plotID"))

#plot stem diameters
veg %>%
  filter(plotID == "WREF_085") %>%                    #looking at only one plot
  ggplot(aes(x = adjEasting, y = adjNorthing)) +
  geom_point(aes(size = stemDiameter/100), shape = 1)

#plot stem diameters
veg %>%
  filter(plotID == "WREF_085") %>%                    #looking at only one plot
  ggplot(aes(x = height)) +
  geom_bar() + 
  labs(title = "histogram of tree heights")
```

```{r download remote sensing data, message = FALSE}

#download AOP data
byTileAOP(dpID = "DP3.30015.001", site = "WREF", year = "2017", # canopy height model
          easting = 580000, northing = 5075000, savepath = data_dir, check.size = F)  #here we're just using one value for E+N, but you can use a vector for E+N as well

```

```{r look at AOP data}

#load: canopy height model
chm <- raster(paste0(data_dir, "/DP3.30015.001/2017/FullSite/D16/2017_WREF_1/L3/DiscreteLidar/CanopyHeightModelGtif/NEON_D16_WREF_DP3_580000_5075000_CHM.tif"))

#basic plot
plot(chm, col = topo.colors(6))

#plot using ggplot
library(viridis)

#write function to convert raster to ggplot-friendly format
raster_to_pts_gg <- function(raster) {
  spdf <- as(raster, "SpatialPixelsDataFrame")
  df <- as.data.frame(spdf)
  colnames(df) <- c("value", "x", "y")
  df
}

chm_gg <- raster_to_pts_gg(chm)

ggplot() + 
  geom_raster(data = chm_gg, aes(x = x, y = y, fill = value)) + 
  scale_fill_viridis() + 
  #remove default plotting background
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

```

```{r extract canopy height model points to }
## extract values of CHM to points in veg data frame
veg_sub <- veg %>%
  filter(!is.na(adjEasting) & !is.na(adjNorthing)) 

veg_xy <- veg_sub %>%
  select(adjEasting, adjNorthing)

#convert veg points to spatial points data frame
coordinates(veg_xy) <- ~adjEasting + adjNorthing
veg_sp <- SpatialPointsDataFrame(coords = veg_xy, data = veg_sub, proj4string = crs(chm))

#extract the values from the canopy height model to the points
model_height <- raster::extract(chm, veg_sp)

#turn extracted points into df to bind
model_height <- data.frame(model_height)

#combine model vals back with original df
veg_full <- bind_cols(veg_sub, model_height) %>%
  mutate(height_diff = height - model_height)

```

```{r inspect model heights vs canopy heights}

#scatterplot
veg_full %>%
  ggplot(aes(x = height, y = model_height)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red")

```
```{r}
#map showing difference

veg_full %>%
  ggplot(aes(x = adjEasting, y = adjNorthing, fill = height_diff)) +
  geom_point(shape = 21) +
  scale_fill_gradient(low = "white", high = "red")

```
```{r}
#3d scatterplot showing difference
plot_ly(veg_full, x = ~adjEasting, y = ~adjNorthing, z = ~height_diff,
        marker = list(color = ~height_diff, colorscale = "Viridis", showscale = TRUE)) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'Easting'),
                     yaxis = list(title = 'Northing'),
                     zaxis = list(title = 'Height model error')),
         annotations = list(
           x = 1.13,
           y = 1.05,
           text = 'm difference',
           xref = 'paper',
           yref = 'paper',
           showarrow = FALSE
         ))

```


```{r}

library(plotly)


# volcano is a numeric matrix that ships with R
p <- plot_ly(z = ~volcano) %>% add_surface()
p

str(volcano)
head(volcano)

```

